---
title: "Project V"
author: "Juan P. Garces"
date: "November 2, 2016"
output: html_document
---
```{r echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library("ggplot2")
library("dplyr")
data("ToothGrowth")
```

### Exploratory Analysis
Below is a datagram showing The Effect of Vitamin C on Tooth Growth in Guinea Pigs
```{r echo=FALSE}
str(ToothGrowth)
```
The Plots below show a comparison between delivery method and dosage.
```{r echo=FALSE}
tGrowth <- ggplot(ToothGrowth, aes(supp,len))+geom_boxplot()+
  labs(x="Supplement Type", y="Tooth Length")
tGrowth

dGrowth <- ggplot(ToothGrowth, aes(as.factor(dose),len))+geom_boxplot()+
  labs(x="Dose in milligrams/day", y="Tooth Length")
dGrowth

bGrowth <- ggplot(ToothGrowth, aes(as.factor(dose),len))+geom_boxplot(aes(fill=supp))+
  labs(x="Dose in milligrams/day", y="Tooth Length")
bGrowth
```


We can appreciate the trend where Orange Juice as a supplement has makes the Guinea Pig's tooth longer, except when a the Dose is 2 milligrams/day. Making ascorbic acid more effective in higher dosage.

##### Means
I created sample means of the amount of growth expected for the two options:
* Guinea pigs who receive a supplement as ascorbic acid (any dosage)
* Guinea pigs who receive 2 mg of the supplement, delivered via orange juice 
```{r}
S1<- filter(ToothGrowth, supp == "VC")
S2<- filter(ToothGrowth, supp == "OJ") # for Later Use
S3<- filter(ToothGrowth, (dose == 2.0 & supp == "OJ")) 
t.test(S1$len, conf.level = .95)
t.test(S3$len, conf.level = .95)
```

### Inferential Analysis
I conducted t-tests to compare mean growth between several different pairs.

#### Comparison between abscorbic acid and orange juice supplement
I test the hypotheses $H_0 : \mu _{VC} = \mu _{OJ}$   versus  $H_a : \mu _{VC} \ne \mu _{OJ}$
```{r}
t <- t.test(len~supp,ToothGrowth)
```
```{r echo=FALSE} 
t 
``` 
The t-statistic of the test was `r t$statistic` and the p-value was `r t$p.value `

SENTENCE 1

#### Comparison between 1mg from orange juice and 2mg dose from orange juice supplement
I test the hypotheses $H_0 : \mu _{1mg} = \mu _{2mg}$   versus  $H_a : \mu _{1mg} < \mu _{2mg}$
```{r}
S4<- filter(S2, dose == 1.0 | dose == 2.0)
t <- t.test(len~dose, S4,alternative = "less")
```
```{r echo=FALSE} 
t 
```
The t-statistic of the test was `r t$statistic` and the p-value was `r t$p.value 

SENTENCE 2

#### Comparison between 1mg of abscorbic acid and 2mg of Abscorbic acid supplement
I test the hypotheses $H_0 : \mu _{1mg} = \mu _{2mg}$   versus  $H_a : \mu _{1mg} < \mu _{2mg}$
```{r}
S5<- filter(S1, dose == 1.0 | dose == 2.0)
t <- t.test(len~dose, S5,alternative= "less")
```
```{r echo=FALSE} 
t 
```
The t-statistic of the test was `r t$statistic` and the p-value was `r t$p.value `

SENTENCE 3

#### Comparison between 1mg abscorbic acid 1mg and orange juice supplement
I test the hypotheses $H_0 : \mu _{1mg VC} = \mu _{1mg OJ}$   versus  $H_a : \mu _{1mg VC} \ne \mu _{1mg OJ}$
```{r}
Both <- filter(ToothGrowth, dose == 1.0)
t <- t.test(len~supp, Both)
```
```{r echo=FALSE} 
t 
```
The t-statistic of the test was `r t$statistic` and the p-value was `r t$p.value` 

SENTENCE 4


#### Comparison between 2mg abscorbic acid and 2mg orange juice supplement
I test the hypotheses $H_0 : \mu _{2mg VC} = \mu _{2mg OJ}$   versus  $H_a : \mu _{2mg VC} \ne \mu _{2mg OJ}$
```{r}
Both <- filter(ToothGrowth, dose == 2.0)
t <- t.test(len~supp, Both)
```
```{r echo=FALSE} 
t 
```
The t-statistic of the test was `r t$statistic` and the p-value was `r t$p.value `

SENTENCE 5





